# Interactive Maps {#interactive}

Resources:

* [Visualization in R](http://remi-daigle.github.io/2016-04-15-UCSB/viz/)
* [Leaflet for R - Introduction](http://rstudio.github.io/leaflet/)
* [mapedit - interactively edit spatial data in R](http://r-spatial.org/r/2017/01/30/mapedit_intro.html)
* [Interactive Viewing of Spatial Objects in R • mapview](https://r-spatial.github.io/mapview/)

## Overview {.objectives}

**Questions**
- How do you generate interactive plots of spatial data?

**Objectives**
- 


## Things You'll Need to Complete this Tutorial {.prereq}

**R Skill Level**: Intermediate - you've got basics of R down.

You'll need ...

We will continue to use the `sf` and `raster` packages and introduce the `leaflet` package in this tutorial.

```{r , message=F, warning=F}
# load packages
library(tidyverse)  # loads dplyr, tidyr, ggplot2 packages
library(sf)         # simple features package - vector
library(raster)     # raster
library(leaflet)    # interactive

# set working directory to data folder
# setwd("pathToDirHere")
```

## States: ggplot2

```{r}
# read in states
states <- read_sf("data/NEON-DS-Site-Layout-Files/US-Boundary-Layers/US-State-Boundaries-Census-2014.shp") %>%
  st_zm() %>%
  mutate(
    water_km2 = (AWATER / (1000*1000)) %>% round(1))

# plot, ggplot
g = ggplot(states) +
  geom_sf(aes(fill = water_km2)) +
  scale_fill_distiller("water_land", palette = "Spectral") +
  ggtitle("Amount of Water by State")
g
```

## States: plotly

```{r, warning=F, message=F}
library(plotly)

ggplotly(g)
```

## States: mapview

```{r, warning=F, message=F}
library(mapview)

mapview(states)
mapview(states, zcol='water_km2', burst='STUSPS') # , burst = TRUE, hide = TRUE) # , burst=T)
```


## States: leaflet

```{r}
library(leaflet)

leaflet(states) %>%
  addTiles() %>%
  addPolygons()
```


### Choropleth

- [Leaflet for R - Choropleths](http://rstudio.github.io/leaflet/choropleths.html)

```{r}
pal <- colorBin("Blues", domain = states$water_km2, bins = 7)


leaflet(states) %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(
    # fill
    fillColor   = ~pal(water_km2),
    fillOpacity = 0.7,
    # line
    dashArray   = "3",
    weight      = 2,
    color       = "white",
    opacity     = 1,
    # interaction
    highlight = highlightOptions(
      weight = 5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE))
```

### Popups and Legend

```{r, warning=F, message=F}
library(htmltools)
library(scales)

labels <- sprintf(
  "<strong>%s</strong><br/> water: %s km<sup>2</sup>",
  states$NAME, comma(states$water_km2)) %>% 
  lapply(HTML)

leaflet(states) %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(
    # fill
    fillColor   = ~pal(water_km2),
    fillOpacity = 0.7,
    # line
    dashArray   = "3",
    weight      = 2,
    color       = "white",
    opacity     = 1,
    # interaction
    highlight = highlightOptions(
      weight = 5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE),
  label = labels,
  labelOptions = labelOptions(
    style = list("font-weight" = "normal", padding = "3px 8px"),
    textsize = "15px",
    direction = "auto")) %>%
  addLegend(
    pal = pal, values = ~water_km2, opacity = 0.7, title = HTML("Water (km<sup>2</sup>)"),
    position = "bottomright")
```

## Pipe Operator

- Help > Keyboard Shortcuts Help. 


## Challenge: Project States and Calculate Area {.challenge}

Use `st_transform()` [USA Contiguous Albers Equal Area Conic: ESRI Projection -- Spatial Reference](http://spatialreference.org/ref/esri/usa-contiguous-albers-equal-area-conic/).

### Answers {.solution} 

- ESRI:102003

```{r}
library(geosphere)
library(units)

# Proj4 of http://spatialreference.org/ref/esri/usa-contiguous-albers-equal-area-conic/
crs_usa = '+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs'

regions = states %>%
  st_transform(crs_usa) %>%
  mutate(
    water_m2 = water_km2 %>% set_units(m^2),
    land_m2  = geometry %>% st_zm() %>% st_area()) %>%
  group_by(region) %>%
  summarize(
    water_m2 = sum(water_m2),
    land_m2  = sum(land_m2)) %>%
  mutate(
    water_land = water_m2 / land_m2)

# table
regions %>%
  st_set_geometry(NULL) %>%
  arrange(desc(water_land))

# plot, ggplot
ggplot(regions) +
  geom_sf(aes(fill = as.numeric(water_land))) +
  scale_fill_distiller("water_land", palette = "Spectral") +
  theme_bw() +
  ggtitle("Ratio of Water to Land (geodesic) for US Regions ")
```

## Key Points {.keypoints}
- Area can be calculated a variety of ways. Geodesic is preferred if starting with geographic coordinates (vs projected).



## Issues

```{r}

```

- `sf::st_is_valid()`

